---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


```{r Loading required libraries, echo = FALSE, results=FALSE, message= FALSE, warning= FALSE}
library(plyr)
library(Hmisc)
library(ggplot2)
```

## Loading and preprocessing the data

```{r FileLoad}
act_data <- read.csv("C:/Users/user/Documents/R/Git_Repos/RepData_PeerAssessment1/activity.csv", stringsAsFactors = FALSE)

act_data$date <- as.Date(act_data$date)

str(act_data)

```


## What is mean total number of steps taken per day?
```{r MTotSteps}
Tot_Num_Step <- aggregate(steps ~ date, data = act_data, sum, na.rm = TRUE)

### Create a histogram of the total number of steps taken each day
hist(Tot_Num_Step$steps, xlab = "Number of Steps" , ylab = , main = "Total number of steps taken per day")

###the mean and median of the total number of steps taken per day
Mean_Total <- as.integer(mean(Tot_Num_Step$steps))

median_Total <- median(Tot_Num_Step$steps)
```

Mean of the total number of steps taken per day is approximately `r Mean_Total`  
Median of the total number of steps taken per day is `r median_Total`



## What is the average daily activity pattern?
```{r AvgSteps}
##time series plot of the 5-minute interval and the average number of steps taken, averaged across all days (y-axis)
AVG_Num_Step <- aggregate(steps ~ interval, data = act_data, mean, na.rm = TRUE)

with(AVG_Num_Step, plot(interval, steps, type = "l"))

AVG_Num_Step <- AVG_Num_Step[order(-AVG_Num_Step$steps),]
```

Maximum number of steps is `r as.integer(AVG_Num_Step[1,2])` for interval `r AVG_Num_Step[1,1]`



## Imputing missing values
```{r NMissing}
NMissing <- sum(is.na(act_data$steps))
```
Number of rows with missing values is `r NMissing`


The Following part will impute missing data based on the mean for the corresponding 5-minute interval
```{r Imputing}
act_data2 <- ddply(act_data, "interval", mutate, imputed_steps = impute(steps, mean))

Tot_Num_Step2 <- aggregate(imputed_steps ~ date, data = act_data2, sum, na.rm = TRUE)

hist(Tot_Num_Step2$imputed_steps, main = "Total number of steps taken each day", xlab = "Number of Steps")

###Total Number of steps before and after imputation
Tot_Num_Step_B <- as.integer(sum(Tot_Num_Step$steps))

Tot_Num_Step_A <- as.integer(sum(Tot_Num_Step2$imputed_steps))
```

Total Number of steps before imputation is `r Tot_Num_Step_B`  
Total Number of steps after imputation is  `r Tot_Num_Step_A`


## Are there differences in activity patterns between weekdays and weekends?
```{r}
##activity patterns between weekdays and weekends
act_data2$Weekdays <- weekdays(act_data2$date)

act_data2$wrkday <- with(act_data2, ifelse(Weekdays== "Saturday","Weekend", ifelse(Weekdays== "Sunday","Weekend", "workday")))

Mean_Num_Step <- aggregate(imputed_steps ~ wrkday + interval, data = act_data2, mean, na.rm = TRUE)

x <- ggplot(Mean_Num_Step,aes(interval,imputed_steps))
x+ geom_line() + facet_grid(.~wrkday)
```



